Fig3

### 2020-03-07 analysis/figure-3.Rmd
better fig 2 and 3

source("../R/themes.R") 
source("../R/functions.R")
source("../R/icons.R")
source("../R/wrangledata.R")

@@

```{r data}
# pca
charpca <- subsetmakepca("pituitary", charlevels, sexlevels)	
charfviz <- makefvizdf("pituitary", charlevels, sexlevels)

@@

@@

@@

abcd <- plot_grid(a,b,c,d, labels = c("b", "c", "d", "e"), ncol = 2, label_size = 8)

expdesign <- png::readPNG("../figures/images/parentalstages_fig1a.png")
expdesign <- ggdraw() +  draw_image(expdesign, scale = 1)

abcde <- plot_grid(expdesign, abcd, nrow = 2, labels = c("a", "b"), label_size = 8, rel_heights = c(0.3,1))
abcde

@@

a <- ggplot(PRLvsd2, aes(x = sex, y = counts, fill = sex))  +
  geom_boxplot() +
  labs(x = "Sex", y = "Prolactin expression")+
  theme_B3() + theme(legend.position = "none") +
  scale_fill_manual(values = sexcolors) +
  geom_hline(yintercept=18, linetype="dashed", color = "black")

b <- ggplot(PRLvsd3, aes(x = sex, y = counts, fill = sex, color = hiloPRL))  +
  geom_boxplot() +
  labs(x = "Prolactin, binned into `lo` and `hi`", y = "Prolactin expression")+
  theme_B3() + theme(legend.position = "none") +
  scale_fill_manual(values = allcolors) +
  scale_color_manual(values = allcolors)

plot_grid(a,b, nrow = 2, rel_widths = c(1,1))

write.csv(PRLvsd3,"../results/PRLvsd.csv", row.names = F)
```

>>>>

source("../R/themes.R") 
source("../R/functions.R")
source("../R/wrangledata.R")

@@

```{r PRLfoo}
# pca
charpca <- subsetmakepca("pituitary", charlevels, sexlevels)	
charfviz <- makefvizdf("pituitary", charlevels, sexlevels)

@@

## Internal versus external

```{r hypothesis}
DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data
DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
DEG_pathfiles <- paste0(DEG_path, DEG_files)
#DEG_files

allDEG2 <- DEG_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name") %>% 
  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/hypothesis/'), "[", 2))  %>% 
  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
  mutate(comparison = paste(down,up, sep = "_")) %>%
  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
head(allDEG2)

allDEG2$tissue <- factor(allDEG2$tissue, levels = tissuelevel)

allDEG2$comparison <- factor(allDEG2$comparison, levels = c("eggs_chicks", "lo_hi"))
allDEG2 <- allDEG2 %>% mutate(comparison = fct_recode(comparison, "lo vs. hi PRL   " = "lo_hi",
                                                      "eggs vs. chicks" = "eggs_chicks"))
allDEG2$direction <- factor(allDEG2$direction, levels = c("eggs", "chicks", "lo", "hi"))

@@

expdesign <- png::readPNG("../figures/images/fig_fig3a.png")
expdesign <- ggdraw() +  draw_image(expdesign, scale = 1)

@@

bcde <- plot_grid(a,b,c,d, labels = c("b", "c", "d", "e"), ncol = 2, label_size = 8)

makenewbargraph <- function(whichtissue, whichsex,  whichcomparison, lowlim, higherlim){
  p <- allDEG2 %>%
    filter(tissue == whichtissue,
           comparison == whichcomparison,
           sex == whichsex) %>%
    ggplot(aes(x = comparison,  fill = direction)) +
    geom_bar(position = "dodge", drop = FALSE) +
    theme_B3() +
    theme(legend.position = "none")  +
    guides(fill = guide_legend(nrow = 1)) +
    labs( y = whichtissue) +
  geom_text(stat='count', aes(label=..count..), vjust =-0.5, 
           position = position_dodge(width = 1),
           size = 2, color = "black")  + 
      ylim(lowlim, higherlim) +
  scale_fill_manual(values = allcolors, name = "higher in")   + 
    theme(axis.text.x = element_blank()) 
  return(p)
}

b11 <- makenewbargraph("hypothalamus", "female","eggs vs. chicks", 0, 2500) +  labs(subtitle = "females", x = NULL, y = " ") + 
  theme( axis.text = element_blank())
b21 <- makenewbargraph("pituitary", "female","eggs vs. chicks", 0, 2500)  + labs(x = NULL, y = " ") + 
  theme(axis.text = element_blank())
b31 <- makenewbargraph("gonad", "female", "eggs vs. chicks", 0, 2500) + labs(x = "eggs vs. chicks", y = " ")    + 
  theme(axis.text = element_blank())
b112131 <- plot_grid(b11,b21,b31, nrow = 3, rel_heights = c(1.1,1,1.1))

b12 <- makenewbargraph("hypothalamus", "male",  "eggs vs. chicks", 0, 2500) + labs(subtitle = "males", x = NULL)+ 
  theme(axis.title.y = element_blank(), axis.text = element_blank())
b22 <- makenewbargraph("pituitary", "male", "eggs vs. chicks", 0, 2500) + labs(x = NULL)+ 
  theme(axis.title.y = element_blank(), axis.text = element_blank())
b32 <- makenewbargraph("gonad", "male", "eggs vs. chicks", 0, 2500) + labs(x = "eggs vs. chicks")   + 
  theme(axis.title.y = element_blank(), axis.text = element_blank())
b122232 <- plot_grid(b12, b22,b32, nrow = 3, rel_heights = c(1.1,1,1.1))

c11 <- makenewbargraph("hypothalamus", "female", "lo vs. hi PRL   ", 0, 2500) +  labs(subtitle = "females", x = NULL, y = "hypothalamus DEGs") + 
  theme( axis.text.x = element_blank()) 
c21 <- makenewbargraph("pituitary", "female", "lo vs. hi PRL   ", 0, 2500)  + labs(x = NULL, y = "pituitary DEGs")+ 
  theme( axis.text.x = element_blank())
c31 <- makenewbargraph("gonad", "female", "lo vs. hi PRL   ", 0, 2500) + labs(x = "lo vs. hi PRL", y = "gonad DEGs")  + 
  theme(axis.text.x = element_blank()) 
c112131 <- plot_grid(c11,c21,c31, nrow = 3, rel_heights = c(1.1,1,1.1))

c12 <- makenewbargraph("hypothalamus", "male",  "lo vs. hi PRL   ", 0, 2500) + labs(subtitle = "males", x = NULL)+ 
  theme(axis.title.y = element_blank(), axis.text = element_blank())
c22 <- makenewbargraph("pituitary", "male", "lo vs. hi PRL   ", 0, 2500) + labs(x = NULL)+ 
  theme(axis.title.y = element_blank(), axis.text = element_blank())
c32 <- makenewbargraph("gonad", "male", "lo vs. hi PRL   ", 0, 2500) + labs(x = "lo vs. hi PRL")   + 
  theme(axis.title.y = element_blank(), axis.text = element_blank())
c122232 <- plot_grid(c12, c22,c32, nrow = 3, rel_heights = c(1.1,1,1.1))

hypothesisbars <-  plot_grid(c112131, c122232, b112131, b122232, nrow = 1, rel_widths = c(1.3, 1, 1.1, 1),
          labels = c("f", " ", "g", " "), label_size = 8)

dataplots <- plot_grid(bcde, hypothesisbars)

fig3 <- plot_grid(expdesign, dataplots, rel_heights = c(0.25,1), nrow = 2, labels = c(" ", "e"), label_size = 8)
fig3

@@

```{r savefiles}
#write.csv(PRLvsd3,"../results/PRLvsd.csv", row.names = F)
